---
title: "Dealing with missing data- Par1 : Mean imputation"
output:
  html_document: default
  word_document: default
---


```{r}
library(tidyverse)
library(DataExplorer)
library(dplyr)
```

A. Detecting missing values

```{r}
data <- read.csv(file = "https://raw.githubusercontent.com/agconti/kaggle-titanic/master/data/train.csv", header = T, sep = ",")
```

```{r}
str(data)
```

The data contain 891 observations and 12 variables.
Interger and numeric type : PassengerId, Survived, Pclass, Age, SibSp, Parch. Fare
Factor type: Name, Sex, Ticket, Cabin, Embarked.

Description of the data set
PassengerID: Id of passengers
Survived:  (0 = No; 1 = Yes)
Pclass:  Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd)
Name: Name
Sex: Sex of passengers
Age: Age
Sibsp: Number of Siblings/Spouses Aboard
Parch: Number of Parents/Children Aboard
Ticket: Ticket Number
Fare: Passenger Fare (British pound)
Cabin:  Cabin
Embarked: Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton)


We can also plot the summary above using plot_intro function.

```{r}
plot_intro(data)
```

There are 5 discret_columns, and 7 continuous columns, corresponding to the 5 columns of factor type and to the 7 columns of numeric and interger type, repectively.
There are 177 missing values and 714 completed rows. 
There are 10692 observations = 891 rows * 12 varibales =  10692 
There is 1.7% missing observations (177/10692 = 1.655%)
complete rows :80.1% (714/891 = 80.13 %)

If we want to know which column containing missing values

```{r}
plot_missing(data)
```

All the missing values are found in the column age.  In fact, 19.87 % of data in this column is missing.
There are 892 rows. There are 19.87% * 891 = 177 missing row, in agreement with the summary above
Other way to check that number 

```{r}
sum(is.na(data$Age))
```

If we want to see missing values for each column
```{r}
sapply(data,function(x)sum(is.na(x)))
```

```{r}
data %>%
  sapply(function(x) sum(is.na(x)))
``` 

B. Imputaion using mean

One way to deal with missing values is to replace NA row by the mean value of the column.
In this data set, we only have missing values for the age column. One way to deal with this situation is to replace the missing value by the average age. 
As the title may be related to the age, we will calculate the average age for each group of people who have the same title. 
As we have many titles (Mr, Ms, etc...) = many genders, we would assign the average age in function of title.
First, we should know how many titles in this data set
What are theses titles ? what are the average age of each of these groups ?
First, we will create a new column that contains only the title. 
Adding a column title into a data set

How does the name column look like ?
```{r}
data %>% 
  select(Name) %>% 
  head(10)
```
You can see that the title is included in the name. We should remove the title and create a new conlumn containing only the title
```{r}
data <- data %>% 
  mutate(Title = regmatches(Name, regexpr("[A-z]+\\.", Name)))
```

```{r}
data %>% 
  select(Name, Title) %>% 
  head(10)
```

To remove title in the column name
```{r}
data <- data %>%
  mutate(Name = gsub("[A-z]+\\.","",Name))
```
To check
```{r}
data %>% 
  select(Name, Title) %>% 
  head(10)
```

Now, we should place the column title in after the column name
```{r}
data <- data %>% 
  select(PassengerId, Name, Title, Sex, Age, everything())
```

```{r}
head(data, 10)
```

Counting the title and order the output
```{r}
data %>% 
group_by (Title) %>%
count(sort = TRUE)
```

There are 17 groups. Mr., Miss., Mrs. are the most common among titles.
There are some umcommon titles such as:Capt, Countess, Don., Jonkheer., Lady., Mme., Ms., and Sir.

Checking the number of missing rows by title

```{r}
data %>%
  group_by(Title) %>% 
  summarise(missing_total = sum(is.na(Age))) %>% 
  arrange(desc(missing_total))
``` 
The missing values were found at corresponding rows : 119 for Mr, 36 for Miss, 17 for Mrs, 4 for Master, and 1 for Dr.
Because there are only 5 groups that contain missing values. We will calculate the mean age of these 5 groups rather than 17 groups

```{r}
data %>%
  filter(Title %in% c("Mr.","Miss.", "Mrs.", "Master.","Dr.")) %>% 
  group_by(Title) %>% 
  summarise (average = mean(Age, na.rm = TRUE))
```

Miss., 182 rows, the  average age = 21.77
Mr., 517 rows, the  average age = 32.36
Mrs.,125 rows, the  average age= 35.89
Master., 40 rows,the  average age = 4.57

Assign the missing values of the age column by respectable values 
Now, we can replace missing values by the mean of corresponding groups
```{r}
data <- data %>% 
  mutate(Age = case_when(Title == "Mr." & is.na(Age) ~ mean(Age[Title == "Mr."], na.rm = TRUE),
                         Title == "Miss." & is.na(Age)~ mean(Age[Title == "Miss."], na.rm = TRUE),
                         Title == "Mrs." & is.na(Age) ~ mean(Age[Title == "Mrs."], na.rm = TRUE),
                         Title == "Master." & is.na(Age) ~ mean(Age[Title == "Master."], na.rm = TRUE), 
                         Title == "Dr." & is.na(Age) ~ mean(Age[Title == "Dr."], na.rm = TRUE),
                         TRUE ~ Age))
```   
         
```{r}
sum(is.na(data$Age))
```

Other way to replace missing value.

train$Age[train$Title == "Mr." & is.na(train$Age)] <- mean(train$Age[train$Title == "Mr."], na.rm = TRUE)

train$Age[train$Title == "Miss." & is.na(train$Age)] <- mean(train$Age[train$Title == "Miss."], na.rm = TRUE)

train$Age[train$Title == "Mrs." & is.na(train$Age)] <- mean(train$Age[train$Title == "Mrs."], na.rm = TRUE)

train$Age[train$Title == "Master." & is.na(train$Age)] <- mean(train$Age[train$Title == "Master."], na.rm = TRUE)


