---
title: "Tidytrick-Replacing string patterns (regrouping categorical variables)"
author: "Minh Trung DANG"
date: "25/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Cheking rows with a specific pattern

```{r}
library(readr)
library(tidyverse)
```

```{r}
incomeUS <- read_csv("adultincome.csv")
```

```{r}
incomeUS %>% 
  group_by(marital.status) %>% 
  summarise(counts = n()) %>% 
  arrange(desc(counts))
```

we can put all the married people in the same group

This varaiable have 7 labels, three of them are:

- Married-AF-spouse: Married armed forces spouse  

- Married-civ-spouse: Married civilian spouse  

- Married-spouse-absent : Married but the spouse was absent

Both levels can be grouped into the group "Married".  

Replace rows containing either "Married-AF-spouse", "Married-civ-spouse", or "Married-spouse-absent" by "Married".


```{r}
patterns <- c("Married-AF-spouse|Married-civ-spouse|Married-spouse-absent")
incomeUS <- incomeUS %>% 
  mutate(marital = stringr::str_replace_all(marital.status, patterns, "Married"))
```

To check the resulting column

```{r}
incomeUS %>% 
  group_by(marital) %>% 
  summarise(counts = n()) %>% 
  arrange(desc(counts))
```


